Operational Feature Selection in Gaussian Mixture Models

### 2016-09-27 paper.tex
update

            \begin{algorithm}
            \caption{Sequential floating forward features selection\label{alg:sffs}}
            {\footnotesize
            \begin{algorithmic}[1]
            \REQUIRE $\Omega,J,\text{maxVarNb}$
            \STATE $\Omega=\emptyset$
            \STATE $F=\text{\{all variables\}}$
            \STATE $k=0$
            \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
            \FORALL{$f_i \in F$}
            \STATE $R_i = J(\{\Omega{(k)} + f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \IF{$R_j \geq J(\Omega^{(k+1)})$}
            \STATE $k=k+1$
            \ELSE
            \STATE $\Omega^{(k+1)} = \{\Omega{(k)} + f_i\}$
            \STATE $k=k+1$
            \STATE $\text{flag}=1$
            \WHILE{$k > 2 \text{ and } \text{flag}=1$}
            \FORALL{$f_i \in \Omega^{(k)}$}
            \STATE $R_i = J(\{\Omega{(k)}\setminus f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \IF{$R_j < J(\Omega^{(k-1)})$}
            \STATE $\Omega^{(k-1)} = \{\Omega{(k)} \setminus f_i\}$
            \STATE $k=k-1$
            \ELSE
            \STATE $\text{flag}=0$
            \ENDIF
            \ENDWHILE
            \ENDIF
            \ENDWHILE
            \RETURN $\Omega$
            \end{algorithmic}
            }
            \end{algorithm}

@@

        Using the formula of the inverse of a block matrix, the following formula expressing $(\boldsymbol{\Sigma}^{(k)})^{-1}$ in function of $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is obtained
        \begin{prop}
        \label{eq:update-inv}
            (Forward update rule for inverse of covariance matrix)
            \begin{equation*}
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                A & v \\
                v^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
            \end{equation*}
        \end{prop}
        where $A = \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} + \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u} \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}}$, $v = - \frac{1}{\alpha} \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\Sigma}^{(k-1)})^{-1}} \mathbf{u}$.

        It means that, to select a variable among one hundred, instead of computing one hundred times $(\boldsymbol{\Sigma}^{(k)})^{-1}$, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed once and then with a few matrix multiplications and sums, the hundred inverses are obtained and a lot of processing time is saved.

        In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes (proof in Appendix~\ref{app:proof-update})
        \begin{prop}
        \label{eq:update-inv-back}
            (Backward update rule for inverse of covariance matrix)
            \begin{equation*}
                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \textcolor[rgb]{0,0.5,0}{\mathbf{A}} - \alpha \mathbf{v} \mathbf{v}^t
            \end{equation*}
        \end{prop}

        A formula can also be deduced for the quadratical term of the decision function. Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained (proof in Appendix~\ref{app:proof-update})
        \begin{prop}
        \label{eq:update-quad}
            (Update rule for quadratical term)
            \begin{align*}
                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}} \\
                &+ \alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2
            \end{align*}
        \end{prop}

        Finally, using the formula of the determinant of a block matrix and after taking the log, a last formula is produced
        \begin{prop}
        \label{eq:update-log}
            (Update rule for logdet)
            \begin{equation*}
                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} + \log \alpha
            \end{equation*}
        \end{prop}

        Now using all update rules, criterion functions can be rewritten as follows
        \begin{align}
        \label{eq:q-update}
            Q(\mathbf{x}) = &- \textcolor[rgb]{0,0.5,0}{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})} \nonumber \\
            &- \alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 \nonumber \\
            &- \textcolor[rgb]{0,0.5,0}{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)} - \log \alpha  + \textcolor[rgb]{0,0.5,0}{2 \log (\pi_c)} + \textcolor[rgb]{0,0.5,0}{k \log(2 \pi)};
        \end{align}

        {\footnotesize
        \begin{align}
        \label{eq:skl-update}
            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \text{Tr} \left((\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_j}^{(k)} + (\boldsymbol{\Sigma}_{c_j}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_i}^{(k)}\right) \nonumber \\
            &+ \textcolor[rgb]{0,0.5,0}{(\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})^t ((\boldsymbol{\Sigma}_{c_i}^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_{c_j}^{(k-1)})^{-1}) (\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})} \nonumber \\
            &+ \alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2 \nonumber \\
            &+ \alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2
            - \textcolor[rgb]{0,0.5,0}{2k} \biggr),
        \end{align}
        }
        with $(\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};

        \begin{align}
        \label{eq:jm-update}
            \text{B}_{ij} = &\textcolor[rgb]{0,0.5,0}{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})} \nonumber \\
            &+ \frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 \nonumber \\
            &+ \textcolor[rgb]{0,0.5,0}{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)} + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right),
        \end{align}
        where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$.

@@

    \begin{algorithm}
    \caption{Sequential floating forward features selection with updates\label{alg:sffs-update}}
    {\footnotesize
    \begin{algorithmic}[1]
    \REQUIRE $\Omega,J,\text{maxVarNb}$
    \STATE $\Omega=\emptyset$
    \STATE $F=\text{\{all variables\}}$
    \STATE $k=0$
    \WHILE{$\text{card}(\Omega) \leq maxVarNb$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\lambda_i$ {\bfseries do} $\lambda_i = \min (\text{EPS\_FLT},\lambda_i)$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
    \FORALL{$f_i \in F$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \min (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega{(k)} + f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j \geq J(\Omega^{(k+1)})$}
    \STATE $k=k+1$
    \ELSE
    \STATE $\Omega^{(k+1)} = \{\Omega{(k)} + f_i\}$
    \STATE $k=k+1$
    \STATE $\text{flag}=1$
    \WHILE{$k > 2 \text{ and } \text{flag}=1$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\lambda_i$ {\bfseries do} $\lambda_i = \min (\text{EPS\_FLT},\lambda_i)$
    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
    \FORALL{$f_i \in \Omega^{(k)}$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \min (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega{(k)}\setminus f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j < J(\Omega^{(k-1)})$}
    \STATE $\Omega^{(k-1)} = \{\Omega{(k)} \setminus f_i\}$
    \STATE $k=k-1$
    \ELSE
    \STATE $\text{flag}=0$
    \ENDIF
    \ENDWHILE
    \ENDIF
    \ENDWHILE
    \RETURN $\Omega$
    \end{algorithmic}
    }
    \end{algorithm}

>>>>

            \begin{algorithm}
            \caption{Sequential floating forward features selection\label{alg:sffs}}
            {\footnotesize
            \begin{algorithmic}[1]
            \REQUIRE $J,\text{maxVarNb}$
            \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
            \STATE $F=\text{\{all variables\}}$
            \STATE $k=0$
            \WHILE{$k \neq \text{maxVarNb}$}
            \FORALL{$f_i \in F$}
            \STATE $R_i = J(\{\Omega_k + f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \STATE $k=k+1$
            \IF{$R_j \geq J(\Omega_k)$}
            \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
            \STATE $\text{flag}=1$
            \WHILE{$k > 2 \text{ and } \text{flag}=1$}
            \FORALL{$f_i \in \Omega_k$}
            \STATE $R_i = J(\{\Omega_k \setminus f_i\})$
            \ENDFOR
            \STATE $j=\text{arg} \max_{i} R_i$
            \IF{$R_j > J(\Omega^{(k-1)})$}
            \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
            \STATE $k=k-1$
            \ELSE
            \STATE $\text{flag}=0$
            \ENDIF
            \ENDWHILE
            \ENDIF
            \ENDWHILE
            \RETURN $\Omega_{\text{maxVarNb}}$
            \end{algorithmic}
            }
            \end{algorithm}

@@

        Using the formula of the inverse of a block matrix, the following formula expressing $(\boldsymbol{\Sigma}^{(k)})^{-1}$ in function of $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is obtained
        \begin{prop}
        \label{eq:update-inv}
            (Forward update rule for inverse of covariance matrix)
            \begin{equation*}
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                A & v \\
                v^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
            \end{equation*}
        \end{prop}
        where $A = (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $v = - \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$ and $ \alpha = \sigma_{kk} - \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}$. Then, $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ is computed only once and this update update formulae is used to obtain all the $(\boldsymbol{\Sigma}^{(k)})^{-1}$ corresponding to all the possible augmented set of a given selection iteration.

        In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes (proof in Appendix~\ref{app:proof-update})
        \begin{prop}
        \label{eq:update-inv-back}
            (Backward update rule for inverse of covariance matrix)
            \begin{equation*}
                (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
            \end{equation*}
        \end{prop}

        A formula can also be deduced for the quadratical term of the decision function. Noting $(\mathbf{x}^{(k)})^t = \left[\begin{array}{cc} (\mathbf{x}^{(k-1)})^t   & x^k \end{array}\right]$, the following proposition is obtained (proof in Appendix~\ref{app:proof-update})
        \begin{prop}
        \label{eq:update-quad}
            (Update rule for quadratical term)
            \begin{align*}
                (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once}\\ \text{per selection step}}} \\
                &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}}
            \end{align*}
        \end{prop}

        Finally, using the formula of the determinant of a block matrix and after taking the log, a last formula is produced
        \begin{prop}
        \label{eq:update-log}
            (Update rule for logdet)
            \begin{equation*}
                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}
            \end{equation*}
        \end{prop}

        Now using all update rules, criterion functions can be rewritten as follows
        \begin{align}
        \label{eq:q-update}
            Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
            &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
            &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}};
        \end{align}

        \begin{align}
        \label{eq:skl-update}
            &\text{SKL}_{ij} = \frac{1}{2} \biggl( \underbrace{\text{Tr} \left((\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_j}^{(k)} + (\boldsymbol{\Sigma}_{c_j}^{(k)})^{-1} \boldsymbol{\Sigma}_{c_i}^{(k)}\right)}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_{c_i}^{(k)} - \boldsymbol{\mu}_{c_j}^{(k)}) )^2 - 2k}_{\substack{\text{computed for} \\ \text{each augmented set}}} \nonumber \\
            &+ \underbrace{(\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})^t ((\boldsymbol{\Sigma}_{c_i}^{(k-1)})^{-1} + (\boldsymbol{\Sigma}_{c_j}^{(k-1)})^{-1}) (\boldsymbol{\mu}_{c_i}^{(k-1)} - \boldsymbol{\mu}_{c_j}^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \biggr),
        \end{align}
        with $(\boldsymbol{\Sigma}_{c_i}^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};

        \begin{align}
        \label{eq:jm-update}
            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once}\\ \text{per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for} \\ \text{each augmented set}}},
        \end{align}
        where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$.

@@

    \begin{algorithm}
    \caption{Sequential floating forward features selection with updates\label{alg:sffs-update}}
    {\footnotesize
    \begin{algorithmic}[1]
    \REQUIRE $J,\text{maxVarNb}$
    \STATE $\Omega=\overbrace{(\emptyset,...,\emptyset)}^{maxVarNb}$
    \STATE $F=\text{\{all variables\}}$
    \STATE $k=0$
    \WHILE{$k \neq \text{maxVarNb}$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
    \STATE Precompute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
    \FORALL{$f_i \in F$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega_k + f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \STATE $k=k+1$
    \IF{$R_j \geq J(\Omega_k)$}
    \STATE $\Omega_k = \{\Omega_{k-1} + f_j\}$
    \STATE $\text{flag}=1$
    \WHILE{$k > 2 \text{ and } \text{flag}=1$}
    \STATE Diagonalize $\boldsymbol{\Sigma}^{(k-1)} = = \mathbf{Q} \boldsymbol{\Lambda} \mathbf{Q}^t$
    \STATE {\bfseries for all} $\Lambda_{i,i}$ {\bfseries do} $\Lambda_{i,i} = \max (\text{EPS\_FLT},\Lambda_{i,i})$
    \STATE Compute {\scriptsize $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$, $(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)}- \boldsymbol{\mu}^{(k-1)})$ and $\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)$} using Propositions \ref{eq:update-inv}, \ref{eq:update-quad} and \ref{eq:update-log}
    \FORALL{$f_i \in \Omega_k$}
    \STATE Compute update constant $\alpha$
    \STATE $\alpha = \max (\text{EPS\_FLT},\alpha)$
    \STATE $R_i = J(\{\Omega_k \setminus f_i\})$ using Equations \ref{eq:q-update}, \ref{eq:skl-update} or \ref{eq:jm-update}
    \ENDFOR
    \STATE $j=\text{arg} \max_{i} R_i$
    \IF{$R_j > J(\Omega_{k-1})$}
    \STATE $\Omega_{k-1} = \{\Omega_k \setminus f_j\}$
    \STATE $k=k-1$
    \ELSE
    \STATE $\text{flag}=0$
    \ENDIF
    \ENDWHILE
    \ENDIF
    \ENDWHILE
    \RETURN $\Omega_{\text{maxVarNb}}$
    \end{algorithmic}
    }
    \end{algorithm}
